---
title: "Shiny exercises"
author: "Jonathan Neimann"
date: "2024-11-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)
```

Problem 1.1
```{r}
library(shiny)
ui <- fluidPage(
  textInput("name", "What's your name?"), # Input field for the user's name
  textOutput("greeting")                 # Output area to display the greeting
)


server <- function(input, output) {
  output$greeting <- renderText({       
    paste0("Hello ", input$name)
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

Problem 1.2
```{r}
ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),
  "then x times 5 is",
  textOutput("product")
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * 5 #did not define input in server logic
  })
}

shinyApp(ui, server)
```
Problem 1.3
```{r}
ui <- fluidPage(
  sliderInput("x", label = "If x is", min = 1, max = 50, value = 30),  
  sliderInput("y", label = "and y is", min = 1, max = 50, value = 5),   #slider for y
  "then, x times y is",
  textOutput("product")                                                 
)

server <- function(input, output, session) {
  output$product <- renderText({ 
    input$x * input$y
  })
}

shinyApp(ui, server)
```
Problem 1.4 

Can use a reactive expression to hold the value of input$x * input$y
```{r}
ui <- fluidPage(
  sliderInput("x", "If x is", min = 1, max = 50, value = 30),
  sliderInput("y", "and y is", min = 1, max = 50, value = 5),
  "then, (x * y) is", textOutput("product"),
  "and, (x * y) + 5 is", textOutput("product_plus5"),
  "and, (x * y) + 10 is", textOutput("product_plus10")
)

server <- function(input, output, session) {
  product <- reactive({ 
    input$x * input$y  # Centralized calculation
  })
  
  output$product <- renderText({ 
    product()  # Use the reactive value
  })
  
  output$product_plus5 <- renderText({ 
    product() + 5  # Add 5 to the reactive value
  })
  
  output$product_plus10 <- renderText({ 
    product() + 10  # Add 10 to the reactive value
  })
}

shinyApp(ui, server)
```
Problem 1.5

```{r}
library(ggplot2)

datasets <- c("economics", "faithfuld", "seals")
ui <- fluidPage(
  selectInput("dataset", "Dataset", choices = datasets),
  verbatimTextOutput("summary"),
  plotOutput("plot") # Bug 1 fixed: changed tableOutput to plotOutput
)

server <- function(input, output, session) {
  dataset <- reactive({
    get(input$dataset, "package:ggplot2")
  })
  
  output$summary <- renderPrint({
    summary(dataset())
  })
  
  output$plot <- renderPlot({
    data <- dataset()
    
    # Bug 2 & 3 fixed: Properly accessing reactive value and creating ggplot
    if (input$dataset == "economics") {
      ggplot(data, aes(x = date, y = unemploy)) +
        geom_line() +
        labs(title = "Unemployment over Time",
             x = "Date",
             y = "Unemployment")
    } else if (input$dataset == "faithfuld") {
      ggplot(data, aes(x = waiting, y = eruptions, fill = density)) +
        geom_tile() +
        scale_fill_viridis_c() +
        labs(title = "Old Faithful Eruption Patterns",
             x = "Waiting Time (minutes)",
             y = "Eruption Time (minutes)")
    } else if (input$dataset == "seals") {
      ggplot(data, aes(x = long, y = lat)) +
        geom_point(alpha = 0.1) +
        labs(title = "Seal Locations",
             x = "Longitude",
             y = "Latitude")
    }
  }, res = 96)
}

shinyApp(ui, server)


```


Create a slider input to select values between 0 and 100 where the interval between each selectable value on the slider is 5. Then, add animation to the input widget so when the user presses play the input widget scrolls through the range automatically.

```{r}
library(shiny)

ui <- fluidPage(
  # Title
  titlePanel("Animated Slider Example"),
  
  # Create the slider input
  sliderInput("slider", "Select Value:", min = 0, max = 100, value = 0, step = 5),
  
  # Create a Play button to trigger animation
  actionButton("play", "Play Animation"),
  
  # Display the selected slider value
  textOutput("slider_value")
)

server <- function(input, output, session) {
  # Reactive values to store animation status and the current value
  animation_running <- reactiveVal(FALSE)
  current_value <- reactiveVal(0)
  
  # Output the current slider value
  output$slider_value <- renderText({
    paste("Current Value: ", input$slider)
  })
  
  # When the play button is clicked, start the animation
  observeEvent(input$play, {
    # Start the animation if it's not already running
    if (!animation_running()) {
      animation_running(TRUE)
      # Run the animation every 200 ms
      for (i in seq(0, 100, by = 5)) {
        invalidateLater(200, session)
        updateSliderInput(session, "slider", value = i)
      }
      # Stop the animation after it finishes
      animation_running(FALSE)
    }
  })
}

shinyApp(ui = ui, server = server)
```

Re-create the Shiny app from Section 2.3.3, this time setting height to 300px and width to 700px. Set the plot “alt” text so that a visually impaired user can tell that its a scatterplot of five random numbers.



```{r}
ui <- fluidPage(
  # Plot Output with description for accessibility
  tags$div(
    # Description for accessibility
    tags$span("A scatterplot of five random numbers on the x-axis (1 to 5), with corresponding values on the y-axis.", 
              style = "font-weight: bold; font-size: 14px;"),
    plotOutput("plot", width = "700px", height = "300px")
  )
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    # Scatterplot of five random numbers
    plot(1:5, y = runif(5), main = "Scatterplot of Random Numbers", 
         xlab = "Index", ylab = "Random Value", pch = 16)
  }, res = 96)
}

shinyApp(ui = ui, server = server)

```

Update the options in the call to renderDataTable() below so that the data is displayed, but all other controls are suppressed (i.e., remove the search, ordering, and filtering commands). You’ll need to read ?renderDataTable and review the options at https://datatables.net/reference/option/.

```{r}
library(DT)

ui <- fluidPage(
  DTOutput("table")  # Use DTOutput() instead of dataTableOutput()
)

server <- function(input, output, session) {
  output$table <- renderDT({
    mtcars
  }, options = list(
    pageLength = 5,
    searching = FALSE,   # Disable the search box
    ordering = FALSE,    # Disable column ordering
    lengthChange = FALSE, # Disable changing the number of rows per page
    filtering = FALSE    # Disable filtering controls
  ))
}

shinyApp(ui = ui, server = server)
```

Problem 3.1

```{r}
ui <- fluidPage(
  textInput("name", "What's your name?"),
  textOutput("greeting")
)

# Server function with the fixed logic (choose one of the corrected versions)
server <- function(input, output, session) {
  output$greeting <- renderText({
    paste0("Hello ", input$name)
  })
}

shinyApp(ui = ui, server = server)
```

Draw the reactive graph for the following server functions:

```{r}
server1 <- function(input, output, session) {
  c <- reactive(input$a + input$b)           # c depends on input$a and input$b
  e <- reactive(c() + input$d)               # e depends on c() and input$d
  output$f <- renderText(e())                # output$f depends on e()
}

server2 <- function(input, output, session) {
  x <- reactive(input$x1 + input$x2 + input$x3)  # x depends on input$x1, input$x2, input$x3
  y <- reactive(input$y1 + input$y2)              # y depends on input$y1 and input$y2
  output$z <- renderText(x() / y())               # output$z depends on x() and y()
}

server3 <- function(input, output, session) {
  d <- reactive(c() ^ input$d)       # d depends on c() and input$d
  a <- reactive(input$a * 10)         # a depends on input$a
  c <- reactive(b() / input$c)       # c depends on b() and input$c
  b <- reactive(a() + input$b)       # b depends on a() and input$b
}
```

Chapter 4

```{r}
dir.create("neiss")
#> Warning in dir.create("neiss"): 'neiss' already exists
download <- function(name) {
  url <- "https://raw.github.com/hadley/mastering-shiny/main/neiss/"
  download.file(paste0(url, name), paste0("neiss/", name), quiet = TRUE)
}
download("injuries.tsv.gz")
download("population.tsv")
download("products.tsv")
```

```{r}
injuries <- vroom::vroom("neiss/injuries.tsv.gz")

```

Problem 4.3

```{r}

ui <- fluidPage(
  titlePanel("Injury Data Summary"),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("prod_code", "Select Product Code:", choices = unique(injuries$product_code)),
      numericInput("num_rows", "Number of Rows to Show:", value = 10, min = 1, step = 1)
    ),
    
    mainPanel(
      tableOutput("diag"),
      tableOutput("body_part"),
      tableOutput("location"),
      plotOutput("age_sex")
    )
  )
)
server <- function(input, output, session) {
  
  # Reactive expression for filtered data based on selected product code
  selected <- reactive({
    filter(injuries, product_code == input$prod_code)
  })
  
  # Reactive expression for summary statistics, grouped by age and sex
  summary_data <- reactive({
    selected_data <- selected()
    summary_table <- selected_data %>%
      group_by(age, sex) %>%
      summarise(injury_count = n(), .groups = "drop") %>%
      mutate(injury_rate = injury_count / sum(injury_count) * 100)
    summary_table
  })
  
  # Display diagnostic counts table, limited by number of rows input
  output$diag <- renderTable({
    head(summary_data() %>% select(diagnosis, injury_count), input$num_rows)
  })
  
  # Display body part counts table, limited by number of rows input
  output$body_part <- renderTable({
    head(summary_data() %>% select(body_part, injury_count), input$num_rows)
  })
  
  # Display location counts table, limited by number of rows input
  output$location <- renderTable({
    head(summary_data() %>% select(location, injury_count), input$num_rows)
  })
  
  # Plot injury summary by age and sex
  output$age_sex <- renderPlot({
    ggplot(summary_data(), aes(x = age, y = injury_count, color = sex)) +
      geom_bar(stat = "identity") +
      labs(title = "Injury Summary by Age and Sex", x = "Age Group", y = "Injury Count")
  })
}

shinyApp(ui = ui, server = server)
```
4.4

```{r}
ui <- fluidPage(
  titlePanel("Narrative Viewer"),
  
  sidebarLayout(
    sidebarPanel(
      actionButton("prev_btn", "Previous Narrative"),
      actionButton("next_btn", "Next Narrative")
    ),
    
    mainPanel(
      textOutput("narrative_text")
    )
  )
)

server <- function(input, output, session) {
  
  # Sample dataset of narratives
  narratives <- reactive({
    c(
      "Narrative 1: The quick brown fox jumps over the lazy dog.",
      "Narrative 2: A journey of a thousand miles begins with a single step.",
      "Narrative 3: To be or not to be, that is the question.",
      "Narrative 4: The only thing we have to fear is fear itself."
    )
  })
  
  # Reactive value to store the current index of the narrative
  current_index <- reactiveVal(1)
  
  # Display the current narrative
  output$narrative_text <- renderText({
    narratives()[current_index()]
  })
  
  # When 'Next' button is clicked, go to the next narrative
  observeEvent(input$next_btn, {
    # Increment the index, loop back to the first narrative if at the end
    new_index <- current_index() + 1
    if (new_index > length(narratives())) {
      new_index <- 1
    }
    current_index(new_index)
  })
  
  # When 'Previous' button is clicked, go to the previous narrative
  observeEvent(input$prev_btn, {
    # Decrement the index, loop back to the last narrative if at the beginning
    new_index <- current_index() - 1
    if (new_index < 1) {
      new_index <- length(narratives())
    }
    current_index(new_index)
  })
}
```

